---
title: "SimPaths results"
format: html
execute:
  echo: false
---

```{r setup}
#| include: false
#| echo: false

library(data.table, quietly = TRUE)
library(purrr, quietly = TRUE)
library(ggpubr, quietly = TRUE)

here::i_am("simpaths-results.qmd")
```

```{r data}
levels_raw <- fread(here::here("data/simpaths_output/cohort_levels.csv"))
effects_raw <- fread(here::here("data/simpaths_output/cohort_effects.csv"))
```

```{r summary-data}
summary_vars <- c(
    "emp_rate",
    "mean_mcs",
    "mean_mcscase45",
    "mean_mhcase",
    "placed_rate",
    "elig_rate"
)

# Median and 95% interval across runs; with a single run the interval collapses to the line
summarise_runs <- function(x) x[, c(
    list(runs = .N, n = median(n)),
    map(.SD, \(v) median(v, na.rm = TRUE)) |> setNames(summary_vars),
    map(.SD, \(v) quantile(v, probs = 0.025, na.rm = TRUE)) |> setNames(paste0(summary_vars, "_l")),
    map(.SD, \(v) quantile(v, probs = 0.975, na.rm = TRUE)) |> setNames(paste0(summary_vars, "_u"))
),
    by = c("scenario", "timescale", "t"),
    .SDcols = summary_vars
]

data <- summarise_runs(levels_raw)
effects <- summarise_runs(effects_raw)
```

```{r plot-layers}
scenarios <- sort(unique(levels_raw$scenario))
scenario_colours <- setNames(
  scales::brewer_pal(palette = "Set1")(max(3, length(scenarios)))[seq_along(scenarios)],
  scenarios
)

common_layers <- list(
  geom_line(aes(color = scenario, linetype = scenario)),
  geom_point(aes(color = scenario, shape = scenario)),
  geom_ribbon(aes(fill = scenario), alpha = 0.05),
  scale_color_manual(values = scenario_colours),
  scale_fill_manual(values = scenario_colours),
  scale_shape_manual(values = setNames(rep_len(c(16, 17, 15, 3), length(scenarios)), scenarios)),
  scale_linetype_manual(values = setNames(rep_len(c("solid", "dashed", "dotted", "dotdash"), length(scenarios)), scenarios)),
  scale_x_continuous(breaks = scales::breaks_pretty()),
  theme_pubr()
)

# k = 0 is the first year a person is eligible in baseline
since_eligible_layers <- list(
  geom_vline(xintercept = -0.5, linetype = "dotted"),
  labs(x = "Years since first eligible")
)

# Effects in percentage points; no fixed accuracy, so neighbouring breaks never share a label
label_pp <- scales::label_number(scale = 100, suffix = " pp")
```

# Cohort

The cohort includes individuals eligible for supported employment at any point in 2021–2026 of the baseline run, followed from the year before they are first eligible.

```{r cohort-size}
#| eval: false
#| include: false
#| echo: false
data[scenario == "baseline" & timescale == "since_eligible", .(
  `Years since first eligible` = t,
  `People followed` = n
)] |>
  knitr::kable()
```

# Levels by years since first eligible

## Employment rate

```{r levels-employment}
#| fig-width: 6
#| fig-height: 4
data[timescale == "since_eligible"] |>
  ggplot() +
  aes(x = t, y = emp_rate, ymin = emp_rate_l, ymax = emp_rate_u) +
  labs(title = "Employment rate in the cohort", y = "Employment rate") +
  scale_y_continuous(labels = scales::label_percent()) +
  common_layers +
  since_eligible_layers
```

## Average MCS

```{r levels-mcs}
#| fig-width: 6
#| fig-height: 4
data[timescale == "since_eligible"] |>
  ggplot() +
  aes(x = t, y = mean_mcs, ymin = mean_mcs_l, ymax = mean_mcs_u) +
  labs(title = "Mean MCS in the cohort", y = "Mean MCS score (0-100)") +
  scale_y_continuous(labels = scales::label_number()) +
  common_layers +
  since_eligible_layers
```

## MCS caseness

```{r levels-mcs-caseness}
#| fig-width: 6
#| fig-height: 4
data[timescale == "since_eligible"] |>
  ggplot() +
  aes(x = t, y = mean_mcscase45, ymin = mean_mcscase45_l, ymax = mean_mcscase45_u) +
  labs(title = "Prevalence of MCS caseness (MCS < 45) in the cohort", y = "Prevalence") +
  scale_y_continuous(labels = scales::label_percent()) +
  common_layers +
  since_eligible_layers
```

# Change relative to baseline by years since first eligible

## Employment rate

```{r effects-employment}
#| fig-width: 6
#| fig-height: 4
effects[timescale == "since_eligible"] |>
  ggplot() +
  aes(x = t, y = emp_rate, ymin = emp_rate_l, ymax = emp_rate_u) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Change in employment rate", y = "Difference from baseline") +
  scale_y_continuous(labels = label_pp) +
  common_layers +
  since_eligible_layers
```

## Average MCS

```{r effects-mcs}
#| fig-width: 6
#| fig-height: 4
effects[timescale == "since_eligible"] |>
  ggplot() +
  aes(x = t, y = mean_mcs, ymin = mean_mcs_l, ymax = mean_mcs_u) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Change in mean MCS", y = "Difference from baseline (MCS points)") +
  scale_y_continuous(labels = scales::label_number()) +
  common_layers +
  since_eligible_layers
```

## MCS caseness

```{r effects-mcs-caseness}
#| fig-width: 6
#| fig-height: 4
effects[timescale == "since_eligible"] |>
  ggplot() +
  aes(x = t, y = mean_mcscase45, ymin = mean_mcscase45_l, ymax = mean_mcscase45_u) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Change in MCS caseness (MCS < 45)", y = "Difference from baseline") +
  scale_y_continuous(labels = label_pp) +
  common_layers +
  since_eligible_layers
```

# Change relative to baseline in 2026

People who have been eligible by 2026, in 2026.

```{r change-mcs}
#| fig-width: 5
#| fig-height: 3
effects |>
  _[timescale == "calendar"] |>
  _[t == 2026] |>
  ggplot() +
  aes(x = mean_mcscase45, xmin = mean_mcscase45_l, xmax = mean_mcscase45_u, y = scenario) +
  geom_pointrange() +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "Change in MCS caseness", y = NULL) +
  scale_x_continuous(labels = label_pp) +
  theme_pubr()
```

```{r change-empl}
#| fig-width: 5
#| fig-height: 3
effects |>
  _[timescale == "calendar"] |>
  _[t == 2026] |>
  ggplot() +
  aes(x = emp_rate, xmin = emp_rate_l, xmax = emp_rate_u, y = scenario) +
  geom_pointrange() +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "Change in employment rate", y = NULL) +
  scale_x_continuous(labels = label_pp) +
  theme_pubr()
```

# Programme reach
```{r reach}
#| fig-width: 6
#| fig-height: 4
#| include: false
#| eval: false
#| echo: false

# Share of the cohort placed by the programme, and share eligible under each scenario's own run. SimPaths draws UC receipt from its global random number generator, so eligibility, and with it who gets placed, is reshuffled between runs; a placement rate well below the eligible rate in baseline dilutes the youth guarantee effects above.

data[timescale == "since_eligible" & scenario != "baseline"] |>
  melt(
    id.vars = c("scenario", "t"),
    measure.vars = c("placed_rate", "elig_rate"),
    variable.name = "measure"
  ) |>
  _[, measure := fifelse(measure == "placed_rate", "Placed", "Eligible in own run")] |>
  ggplot() +
  aes(x = t, y = value, color = scenario, linetype = measure) +
  geom_line() +
  geom_point() +
  labs(title = "Programme reach in the cohort", y = "Share of cohort", linetype = NULL) +
  scale_color_manual(values = scenario_colours) +
  scale_x_continuous(breaks = scales::breaks_pretty()) +
  scale_y_continuous(labels = scales::label_percent()) +
  theme_pubr() +
  theme(legend.box = "vertical") +
  since_eligible_layers
```
